---
title: "Census Data for Final Project"
author: "Kylie Wheelock Riley"
date: "2/27/2021"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidycensus)
library(tidyverse)
library(dplyr)

census_api_key("545c7215e59e244a022aa2f8ba7c4fd66ab3649a")
```

Getting the population for each CA county from 2010 census
```{r County Population 2010}
county_pop = get_decennial(geography = "county",
                           variables = "P001001",
                           state = "CA"
                           )
county_pop %>%
  ggplot(aes(x = value, y = reorder(NAME, value))) + 
  geom_point()
 
```
Los Angeles county is the most populous county in California with just under 10 million people, San Diego and Orange County come in second and third, respectively, with around 3 million each. Alpine County is the least populous county. 


Calculating the percent urban and percent rural housing from each CA county from 2010 census data 
```{r Land Use}

##Search for land use/ urban rural vars by pulling codebook for census vars
cdbk2010 = load_variables(2010, "sf1", cache = TRUE)

View(cdbk2010)
##Create a list of the relevant urban/ rural vars
urbanvars = c("H001001", "H002002", "H002005")

##Pull urban/ rural values, move to wide format, and rename census vars
urban_housing = 
  get_decennial(
    geography = "county",
    variables = urbanvars,
    state = "CA"
    ) %>% 
  pivot_wider(
    names_from = "variable", 
    values_from = "value"
    ) %>% 
  dplyr::rename(
        tot_housing = H001001,
        urban_hous =  H002002, 
        rural_hous = H002005 
        ) 
  
##calculate % rural and % urban vars
urban_housing = mutate(urban_housing, 
         pct_urb = urban_hous / tot_housing, 
         pct_rural = rural_hous / tot_housing
         )
  
##plot % urban
urban_housing %>%
  ggplot(aes(x = pct_urb, y = reorder(NAME, pct_urb))) + 
  geom_point()

##plot % rural
urban_housing %>%
  ggplot(aes(x = pct_rural, y = reorder(NAME, pct_rural))) + 
  geom_point()
```

Viewing household income from 
```{r Household Income using ACS}
household_income = get_acs(geography = "county",
                           variables = "B19013_001",
                           state = "CA"
                           )
household_income %>%
  ggplot(aes(x = estimate, y = reorder(NAME, estimate))) + 
  geom_point()

```

Creating census dataset with all relevant variables 